Torrential Twitter? Measuring the Severity of Harassment when Canadian Female Politicians Tweet about Climate Change

### 2024-02-03 Outputs/Paper/Publication_version/paper_for_pub.qmd
small edits

---
title: "Torrential Twitter? Measuring the Severity of Harassment when Canadian Female Politicians Tweet about Climate Change"
author: "Inessa De Angelis^[Faculty of Information, University of Toronto, Inessa.DeAngelis@mail.utoronto.ca]"
thanks: "Code and select data available at: https://github.com/InessaDeAngelis/torrential_twitter_bell_conf "
date: "today"
date-format: "long"
abstract: "Research has shown that women in politics face considerable harassment. Known as “gendertrolling,” this harassment is part of a global trend of violence against women in politics. This study investigates the severity of harassment three Canadian female politicians receive when Tweeting about climate change and the types of accounts behind the replies. Results from the quantitative content analysis reveal that 86% of Tweets contained some form of harassment. I then built a Bayesian hierarchical model to investigate the relationship between the severity of harassment and the type of account that a user has. The results from my model suggest that despite differences in political affiliation and status, the three politicians are equally impacted by harassment when Tweeting about climate change. *This paper contains language and themes that some readers may find offensive.*"
format: pdf
toc: true
number-sections: true
bibliography: publication_references.bib
csl: apa.csl
---

@@

"F\*CK OFF WITH YOUR CLIMATE SCAM .. nazi traitor b!tch," reads a reply sent to a Tweet about climate change that originated on the account of Canada's Deputy Prime Minister (DPM) and Minister of Finance, the Honourable Chrystia Freeland. For DPM Freeland and women across all political parties and levels of government in Canada and other democratic nations, coping with online and offline harassment is a reality of serving as an elected official [@southern; @silencingwomen]. Known as "gendertrolling," this harassment is part of a global trend of violence against women [@mantilla2013; @wagner]. Gender-based harassment on Twitter is well documented in academic literature and inexplicably pro-environmental behaviors are often viewed as "feminine" [@Citron; @mediatingmisogyny; @Anshelm]. Female politicians who adopt strong policy stances are more likely to face online misogyny and harassment because it is not consistent with existing gender stereotypes and notions of power and poses a threat to male climate deniers [@courtemanche; @rheault; @degeus].

@@

Type of account is defined as the dynamics of relationships between followers, followings, and lists and will be determined by analyzing the accounts' bio, profile image, and other characteristics [@singh; @uddin]. Types of accounts include:

@@

By employing the theory of deliberative democracy, this paper analyzes responses to Tweets sent by female politicians which explicitly discuss climate change and environmental policy to evaluate the extent at which they fulfill the characteristics of healthy democratic discourse. Often paired with their work on the public sphere, Jürgen Habermas and Richard Sennet are credited for publishing founding research on the theory of deliberative democracy [@kies2010]. Recently, communication and media studies scholars have reconceptualized the formation of the public sphere, evaluating the role of platforms with promised democratizing qualities, disinformation, and how these mediums impact the original notion of deliberative democracy [@bruns2023].

@@

All Twitter data was collected throughout the month of October 2023 using the statistical programming software R [@r] and functions from `voson.tcn` [@gertzel2021] and the `tidyverse` [@rTidyverse]. Paid access to Twitter's Basic v2 Application Programming Interface (API) endpoint was obtained before using the advanced search feature to determine which Tweets I wanted to download via the API. My sample includes any original Tweets written in English (no quote or Retweets) that originated on the accounts of DPM Freeland (\@cafreeland), MP May (\@elizabethmay), and MP Collins (\@laurel_bc) that were posted between September 1, 2022, and October 31, 2023. Tweets also had to the contain the word "climate,” "environment,” and/or "environmental" (as part of the text of the Tweet or as a hashtag). For threaded Tweets sent by the politicians, the first Tweet of the thread had to contain at least one of the key words and the overall focus of the thread had to be on climate change and/or environmental policy. For Tweet threads that mentioned one of the key words in later Tweets and in passing, these Tweets and replies were omitted due to their lack of relevance. Key words were also checked for applicability, given the use of words such as "climate" and "environment" to reference geopolitics and economic situations and not actual climate or environmental policy. Within the parameters of the sample, DPM Freeland's Tweets were from October 2022 to October 2023 and had either the key word "climate" or "environment". MP May's Tweets were from September 2022 to October 2023 and included the words "climate," "environment," and "environmental". Lastly, MP Collins' Tweets were from October 2022 to October 2023 and included the words "climate," "environment," and "environmental."

@@

The raw datasets underwent multiple cleaning procedures and tests before being ready to manually code. Cleaning refers to the process of preparing a raw dataset for further analysis, ensuring basic aspects like each Tweet and corresponding account metadata has its own column, row, and cell, before moving onto project-specific preparations [@tellingstories]. Cleaning serves as a useful process for identifying anomalies in the dataset, such as mis-matched Tweets and account metadata. The `voson.tcn` package provided several columns of Twitter metadata that was not relevant to the objectives of this paper, so I eliminated those and focused on columns such as the text of the Tweet, the user's location, bio, profile picture, and whether they paid to be verified with a blue checkmark through Twitter's subscription program. All raw and coded data cleaning was done using R [@r], utilizing functions from the `tidyverse` [@rTidyverse] and `janitor` [@rJanitor]. The original Tweets and replies sent by the female politicians in response to their own thread (known as a threaded Tweet) were omitted during the cleaning process. The few replies sent in languages other than English, including French, Ukrainian, German, and Arabic were removed. Testing of the combined and cleaned datasets was conducted using `validate` [@validate] before being manually coded. The final number of Tweets left following these cleaning procedures are displayed in @tbl-breakdown, which highlights that DPM Freeland is overrepresented in my sample.

@@

After manually coding the 4,139 Tweets using my Codebook (see [Appendix]), I proceeded to the second round of cleaning. I eliminated columns of data which were useful for determining the type of account but not relevant for the next steps of my analysis. For my severity of harassment data, I updated value 1 to "Positive," 2 to "Neutral," 3 to "Questioning authority," 4 to "Name-calling/Gender insults," 5 to "Vicious language," 6 to "Credible threats," and 7 to "Hate speech" based on my Codebook. Furthermore, for my type of account data, I changed value 1 to "Personal," 2 to "Professional," 3 to "Bots," 4 to "Spammers," 5 to "Anonymous," and 6 to "Suspended/deleted" based on my Codebook. I tested the cleaned data using `validate` [@validate], primarily validating variable types and ensuring the number of rows in the dataset matches the number of Tweets I coded. I then analyzed and visualized my data using functions from `knitr` [@rknitr], `kableExtra` [@rKableExtra], `Formattable` [@rFormattable], and `ggplot2` [@rGgplot2].

@@

Twitter's ever-changing affordances and focuses as a platform (such as promoting "free speech") directly and indirectly impact the severity of harassment and types of accounts deemed acceptable to keep on the platform. Many of the third-party tools, such as Botometer [@botometer], which previously assisted computational social science researchers in determining types of accounts have been rendered unusable with Twitter API changes, limiting my ability to determine which accounts in my study are bots. Consequently, the number of accounts categorized as bots in my study is likely smaller than the actual number of bots. It is impossible to know on a granular level how every change to Twitter affordances impacts each harassing Tweet and account on the platform, but can generally inform hypotheses and patterns established when analyzing the data.

Given the in-progress nature of this study, only one coder has coded all the Tweets so far, meaning there is no intercoder reliability. Following the methods outlined by @krippendorff, two additional coders will code a subset of Tweets, before statistically measuring intercoder agreement using `krippendorffsalpha` [@krippendorffsalpha].

@@

In this paper, I examined Tweets explicitly discussing climate change and environmental policy sent to DPM Freeland, MP May, and MP Collins to measure the severity of harassment on a seven-point scale and type of accounts behind the Tweets. My results reveal that over 85% of Tweets sent to the three politicians contained some level of harassment. My model identified and predicted nearly no quantitative differences between Tweets sent to all three female politicians. Spammer and Anonymous accounts were most frequently associated with engaging with the female politicians on Twitter and being responsible for sending the more serve forms of harassment identified. The Tweets analyzed for this study generally matched the criteria of what constitutes healthy democratic discourse, with users with multiple, diverse viewpoints weighing in on the issue of climate change. In future iterations of this study, attention will be turned toward ensuring intercoder reliability, gleaning a better understanding of the frequency of words used in harassing Tweets by doing topic modeling using Latent Dirichlet Allocation, and using LLMs to detect harassment. It is crucial to ensure we continue studying and discussing the intersection of climate change denialism and misogyny, because if we cannot mediate the risks of gendertrolling at the highest echelons of power, then female politicians may self-censor and risk everyone's ability to engage in respectful discourse surrounding climate change [@mediatingmisogyny].

# Declaration of Conflicting Interests {.unnumbered}
The author declares no potential conflicts of interest with respect to the research, authorship, and/or publication of this article.

# Acknowledgments {.unnumbered}
I am grateful for Rohan Alexander's helpful suggestions and feedback on drafts of this paper and for Elizabeth Dubois' guidance during the early conception of this study. Thank you to Sarah Xie and Allyson Cui for their research assistance.

# Funding {.unnumbered}
The author discloses receipt of the following financial support for the research, authorship, and/or publication of this article: This study was funded by the Social Sciences and Humanities Research Council of Canada, Canada Graduate Scholarship (id: 766-2022-0719) and the Government of Ontario, *2023-2024 Ontario Graduate Scholarship*.

# Appendix {.unnumbered}

  - My codebook is available [here](https://github.com/InessaDeAngelis/torrential_twitter_bell_conf/blob/main/Inputs/Codebook.pdf)

>>>>

---
title: "Torrential Twitter? Measuring the Severity of Harassment when Canadian Female Politicians Tweet about Climate Change"
author: "Inessa De Angelis^[Faculty of Information, University of Toronto, Inessa.DeAngelis@mail.utoronto.ca]"
thanks: "Code and select data available at: https://github.com/InessaDeAngelis/torrential_twitter_bell_conf "
date: "today"
date-format: "long"
abstract: "Research has shown that women in politics face considerable online harassment know as “gendertrolling.” This study investigates the severity of harassment three Canadian female politicians receive when Tweeting about climate change and the types of accounts behind the replies. All replies to climate and environmental focused Tweets posted by the politicians from September 2022 to October 2023 were collected for quantitative content analysis. Results from my content analysis reveal that 86% of Tweets contained some form of harassment. I then built a Bayesian hierarchical model to investigate the relationship between the severity of harassment and the type of account that a user has. The results from my model suggest that despite differences in political affiliation and status, the three politicians are equally impacted by harassment when Tweeting about climate change. *This paper contains language and themes that some readers may find offensive.*"
format: pdf
toc: true
number-sections: true
bibliography: publication_references.bib
csl: apa.csl
---

@@

"F\*CK OFF WITH YOUR CLIMATE SCAM .. nazi traitor b!tch," reads a reply sent to a Tweet[^1] about climate change that originated on the account of Canada's Deputy Prime Minister (DPM) and Minister of Finance, the Honourable Chrystia Freeland. For DPM Freeland and women across all political parties and levels of government in Canada and other democratic nations, coping with online and offline harassment is a reality of serving as an elected official [@southern; @silencingwomen]. Known as "gendertrolling," this harassment is part of a global trend of violence against women [@mantilla2013; @wagner]. Gender-based harassment on Twitter is well documented in academic literature and inexplicably pro-environmental behaviors are often viewed as "feminine" [@Citron; @mediatingmisogyny; @Anshelm]. Female politicians who adopt strong policy stances are more likely to face online misogyny and harassment because it is not consistent with existing gender stereotypes and notions of power and poses a threat to male climate deniers [@courtemanche; @rheault; @degeus].

[^1]: In July 2023, CEO Elon Musk announced that Twitter would be renamed to X. This paper will continue to call the platform Twitter.

@@

Type of account is defined as the dynamics of relationships between followers, followings, and lists and will be determined by analysing the accounts' bio, profile image, and other characteristics [@singh; @uddin]. Types of accounts include:

@@

By employing the theory of deliberative democracy, this paper analyses responses to Tweets sent by female politicians which explicitly discuss climate change and environmental policy to evaluate the extent at which they fulfill the characteristics of healthy democratic discourse. Often paired with their work on the public sphere, Jürgen Habermas and Richard Sennet are credited for publishing founding research on the theory of deliberative democracy [@kies2010]. Recently, communication and media studies scholars have reconceptualized the formation of the public sphere, evaluating the role of platforms with promised democratizing qualities, disinformation, and how these mediums impact the original notion of deliberative democracy [@bruns2023].

@@

All Twitter data was collected throughout the month of October 2023 using the statistical programming software R [@r] and functions from `voson.tcn` [@gertzel2021] and the `tidyverse` [@rTidyverse]. Paid access to Twitter's Basic v2 Application Programming Interface (API) endpoint was obtained before using the advanced search feature to determine which Tweets I wanted to download via the API. My sample includes any original Tweets written in English (no quote or Retweets) that originated on the accounts of DPM Freeland (\@cafreeland), MP May (\@elizabethmay), and MP Collins (\@laurel_bc) that were posted between September 1, 2022, and October 31, 2023. These three politicians were selected based on their pro-climate change stances, evidence of harassing Tweets, and diversity in regional representation and political affiliation. Only 6% of female MPs identify as a visible minority or Indigenous and do not frequently discuss climate change on Twitter, which is why my sample contains all white MPs [@hoc]. Tweets also had to the contain the word "climate," "environment," and/or "environmental" (as part of the text of the Tweet or as a hashtag). For threaded Tweets sent by the politicians, the first Tweet of the thread had to contain at least one of the key words and the overall focus of the thread had to be on climate change and/or environmental policy. For Tweet threads that mentioned one of the key words in later Tweets and in passing, these Tweets and replies were omitted due to their lack of relevance. Key words were also checked for applicability, given the use of words such as "climate" and "environment" to reference geopolitics and economic situations and not actual climate or environmental policy. Within the parameters of the sample, DPM Freeland's Tweets were from October 2022 to October 2023 and had either the key word "climate" or "environment". MP May's Tweets were from September 2022 to October 2023 and included the words "climate," "environment," and "environmental". Lastly, MP Collins' Tweets were from October 2022 to October 2023 and included the words "climate," "environment," and "environmental."

@@

The raw datasets underwent multiple cleaning procedures and tests before being ready to manually code. Cleaning refers to the process of preparing a raw dataset for further analysis, ensuring basic aspects like each Tweet and corresponding account metadata has its own column, row, and cell, before moving onto project-specific preparations [@tellingstories]. Cleaning serves as a useful process for identifying anomalies in the dataset, such as mis-matched Tweets and account metadata. I eliminated columns of metadata unrelated to this study and focused on columns such as the text of the Tweet, the user's location, bio, profile picture, and whether they paid to be verified with a blue checkmark through Twitter's subscription program. All raw and coded data cleaning was done using R [@r], utilizing functions from the `tidyverse` [@rTidyverse] and `janitor` [@rJanitor]. The original Tweets and replies sent by the female politicians in response to their own thread (known as a threaded Tweet) were omitted during the cleaning process. The few replies sent in languages other than English, including French, Ukrainian, German, and Arabic were removed. Testing of the combined and cleaned datasets was conducted using `validate` [@validate] before being manually coded. The final number of Tweets left following these cleaning procedures are displayed in @tbl-breakdown, which highlights that DPM Freeland is overrepresented in my sample.

@@

After manually coding the 4,139 Tweets using my Codebook (see [Appendix]), I proceeded to the second round of cleaning. I eliminated columns of data which were useful for determining the type of account but not relevant for the next steps of my analysis. For my severity of harassment data, I updated value 1 to "Positive," 2 to "Neutral," 3 to "Questioning authority," 4 to "Name-calling/Gender insults," 5 to "Vicious language," 6 to "Credible threats," and 7 to "Hate speech" based on my Codebook. Furthermore, for my type of account data, I changed value 1 to "Personal," 2 to "Professional," 3 to "Bots," 4 to "Spammers," 5 to "Anonymous," and 6 to "Suspended/deleted" based on my Codebook. I tested the cleaned data using `validate` [@validate], primarily validating variable types and ensuring the number of rows in the dataset matches the number of Tweets I coded. I then analysed and visualized my data using functions from `knitr` [@rknitr], `kableExtra` [@rKableExtra], `Formattable` [@rFormattable], and `ggplot2` [@rGgplot2].

@@

Twitter's ever-changing affordances and focuses as a platform (such as promoting "free speech") directly and indirectly impact the severity of harassment and types of accounts deemed acceptable to keep on the platform. Many of the third-party tools, such as Botometer [@botometer], which previously assisted computational social science researchers in determining types of accounts have been rendered unusable with Twitter API changes, limiting my ability to determine which accounts in my study are bots. Consequently, the number of accounts categorized as bots in my study is likely smaller than the actual number of bots. It is impossible to know on a granular level how every change to Twitter affordances impacts each harassing Tweet and account on the platform, but can generally inform hypotheses and patterns established when analysing the data.

@@

In this paper, I examined Tweets explicitly discussing climate change and environmental policy sent to DPM Freeland, MP May, and MP Collins to measure the severity of harassment on a seven-point scale and type of accounts behind the Tweets. My results reveal that over 85% of Tweets sent to the three politicians contained some level of harassment. My model identified and predicted nearly no quantitative differences between Tweets sent to all three female politicians. Spammer and Anonymous accounts were most frequently associated with engaging with the female politicians on Twitter and being responsible for sending the more serve forms of harassment identified. The Tweets analysed for this study generally matched the criteria of what constitutes healthy democratic discourse, with users with multiple, diverse viewpoints weighing in on the issue of climate change. In future iterations of this study, attention will be turned toward using LLMs to detect nauanced forms of harassment. It is crucial to ensure we continue studying and discussing the intersection of climate change denialism and misogyny, because if we cannot mediate the risks of gendertrolling at the highest echelons of power, then female politicians may self-censor and risk everyone's ability to engage in respectful discourse surrounding climate change [@mediatingmisogyny].

# Acknowledgments {.unnumbered}
The author(s) thank(s) Rohan Alexander for the helpful suggestions and feedback on drafts of this paper and Elizabeth Dubois for guidance during the early conception of this study. The author(s) also thank(s) Sarah Xie and Allyson Cui for their research assistance.

# Funding {.unnumbered}
The author(s) disclose(s) receipt of the following financial support for the research, authorship, and/or publication of this article: This study was funded by the Social Sciences and Humanities Research Council of Canada (grant ID: 766-2022-0719) and the Government of Ontario, *2023-2024 Ontario Graduate Scholarship*.

# Declaration of Conflicting Interests {.unnumbered}
The author(s) declare(s) that there is no conflict of interest.
